### 2009-05-12 doc/tech_doc/infosys/infosys_technical.tex
Moving more from handbook. Started adding security parts.

\chapter{Design Overview} % (fold)
\label{cha:design_overview}

\chapter{ISIS} % (fold)
\label{cha:isis}

\section{Registration handling} % (fold)
\label{sec:registration_handling}

\subsection{Functionality} % (fold)
\label{sub:isis_functionality}

@@

\subsection{Interface} % (fold)
\label{sub:isis_interface}

@@

This operation is usually called by service which wants to register it's presence in ISIS. This message consists of one or more \textbf{Registration Entry} and at most one \textbf{Registration Header}. The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv). This structure is shown on Figure \ref{fig:RegistrationMessage}.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
\caption{\label{fig:RegistrationMessage}Embedded structure of Registration Message} }
\end{figure}

@@

\section{Peer-to-Peer} % (fold)
\label{sec:peer_to_peer}

\subsection{Functionality} % (fold)
\label{sub:p2p_functionality}

@@

\subsection{Interface} % (fold)
\label{sub:p2p_interface}

@@

@@

\chapter{Service} % (fold)
\label{cha:service}

\section{Information} % (fold)
\label{sec:service_information}

@@

@@

\section{Registration} % (fold)
\label{sec:service_registration}

@@

Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{isis_interface}).
Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).

@@

>>>>

\chapter{Design Overview}
\label{cha:design_overview}

\chapter{ISIS}
\label{cha:isis}

\section{Registration handling}
\label{sec:isis_registration_handling}

\subsection{Functionality}
\label{sub:isis_functionality}

@@

\subsection{Interface} 
\label{sub:isis_interface}

@@

This operation is usually called by service which wants to register it's presence in ISIS. This message consists of one or more \textbf{Registration Entry} and at most one \textbf{Registration Header}. The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv). This structure is shown on Figure \ref{fig:registration_message}.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
\caption{\label{fig:registration_message}Embedded structure of Registration Message} }
\end{figure}

@@

\section{Peer-to-Peer}
\label{sec:isis_peer_to_peer}

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}

%\subsubsection{Peer-to-Peer functions}
%\label{ISIS Peer-to-Peer functions}
%\subsection{Client}
%\label{Client}

%The client part of the information service is for querying the data stored in the database of the ISIS 
%services. This is implemented as a part of the ARCLib and using the ISIS service's interface for the 
%various data queries. For further details see Section \ref{Query data}.

%\section{Operations}
%\label{Operations}
%The different operations are performed by different Information system components. These cases are described below in the dedicated subsections.

@@

\subsection{Interface}
\label{sub:isis_peer_to_peer_interface}

@@

\section{Authorization}
\label{sec:isis_authorization}

\subsection{Client Authorzation}
\label{sub:isis_authorization_client_authorization}

% TODO figure with mutiple service types

To ensure information stored in ISIS cloud can't be tampered and only available to proper clients following authorization framework is implemented. All actions perfored by ISIS clients are divided into three following groups:

\begin{itemize}
 \item Operations initiated by other ISISes in the cloud. Those include:
 \begin{itemize}
  \item Register with Registration Message containing information not about contacting client
  \item RemoveRegistrations with request to remove Registration Message representing not contacting client
  \item Connect % TODO: describe this operation
 \end{itemize}

 Those operations may cause uncontrollable changes in colleccted information and must be granted only for higly trusted entities like ISISes themselves.

 \item Operations initiated by the Services registering to Information System. Those are:
 \begin{itemize}
  \item Register with Registration Message containing information about contacting client
  \item RemoveRegistrations with request to remove Registration Message representing contacting client
 \end{itemize}

 \item Operations which are allowed for any liable client of particular Grid infrastructure.
 \begin{itemize}
  \item Query
  \item GetISISList
 \end{itemize}
\end{itemize}

% subsection client_authorization (end)

\subsection{Information Authorzation}
\label{sub:isis_authorization_information_authorization}

%* Clasification of information
%
% 1. From acccessibility point of view information can be devided
%into 3 groups:
% a. public
% b. available to trusted clients (like broker or scheduler service)
% c. private - available to owner of information object and entities
%trusted by owner.
%Distinction between 2 and 3 is not clear. One may accept that 2 is
%imposed by architecture/deployment of system and 3 is controlled by
%entities which are users of the system.
%
 2. From storage/access location point of view information is divided
%into:
% a. local - stored at endpoint representing resource which is
%described by information. This information is obtainable through
%ALIS interface.
% b. propagated to dedicated services (ISIS) which collect information
%needed for pre-selection and initial lookup of services
%representing resources.
%
%NOTE: Definition of propagation of information does not exactly belong
%to security architectire of information system. But it affects it significantly
%and hence at least minimal set of definitions is needed.
%
%OUT OF BAND NOTE: Those services could provide pre-selection capability.
%
% Sets 1 and 2 may intersect in any way. In first approach to simplify
%implementation 2b may be considered to always belong to 1a.
%
%
%* Implementaion guidelines
%
% To implement set 1 information objects need to have Access Control
%Rules (ACR) assigned (usually used term Access Control List seems to
%be inadequate due to meaning of word List).
%
% To implement set 2 in similar way information objects have Propagation
%Control Rules (PCR) assigned. Because curently information system is made
%of one layer of ISIS PCRs may consist of one boolean property.
%
% Information is structured in tree-like fashion - XML document.
%Access control is defined at node and follows all branches - similar
%to file systems. Propagation control is similar to access control (if
%needed at all). Children nodes can't widen access defined at parent
%nodes. This should decrease amount of ACR to be evaluated.
%
%EXAMPLE: if parent node A allows access only to VO_1 then children
%node B can narrow access to Adminsitrator of VO_1 and can't open
%access for VO_2.
%
% By default all nodes are public and are propagatable. Properties
%assigned to clildren nodes limit access and propagation. Access
%permission to particular information node is evaluated by traversing
%all nodes prom parent to child. At first node which gives negative
%result evaluation is stopped.
%
% Because evaluation of permission may be costly (CPU intensive,
%causing big latency if access to external service is needed, etc.)
%other tricks to reduce cost are needed.
%
% To reduce access control evaluation cost ACRs may/should/must(?)
%be assigned to information nodes in one-to-many way - one ACR per
%multiple information nodes. This grouping may be initiated either
%interactively (by clients explicitely specifying that same ACR is
%applicable to multiple nodes) or automatically (by comapring new
%ACR to already existing).
%
% If information nodes containing ACRs are propagated to ISIS (second
%implementation step) propagation algorithm must take ACR into account.
%For informtion propagation between ISISes this rule also applies.
%EXAMPLE: Information with access open to VO_1 must not be propagated
%to ISIS belonging to VO_2.
%
%
%** Ideas for in-memory XML documents
%
% Because informational documents are normally produced dynamically and
%have short lifetime there seems to be reasonable to keep them in memory.
% Each node with assigned ACR contains attribute (even if that breaks schema)
%containing id of ACR. ACR is defined as XML element (even if that breaks
%schema) located at higher, same or child level of considered node.
% Because informational documents are produced dynamically there are few
%ways for processing ACRs:
% 1. Right after document is generated it is traversed and ACRs applied.
%Parts of document which do not pass ACR evaluation are cut off. ACRs are
%cut off too.
% 2. While calling document generation procedure identity of client is
%supplied and ACRs are applied immediately.
% 3. While calling document generation procedure hooks/callbacks are supplied
%which are used and implement ACR evaluation.
% Combination of 1 and 3 looks most promising because it is flexible and
%keeps information system and policy evaluation code separated.
%
%
%* References
%"Information access control use cases" http://wiki.nordugrid.org/index.php/ARC1/Infosys/Access_Use_cases
%

% subsection information_authorization (end)

% section authorization (end)

@@

\chapter{Service}
\label{cha:service}

\section{Information generation}
\label{sec:service_information_generation}

@@

Because there may be multiple registration processes running in parallel it is important to ensure that implementation RegistrationCollector is thread safe or there are internal locks implemented.

@@

\section{Registration}
\label{sec:service_registration}

@@

Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).

An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
there are two Registrants configured in one HED container for three services. The \textit{Registrant A} is the 
default Registrant and the services configured in the following way:

\begin{itemize}
  \item Service 1: There is no Registrant configured so the default one will register it.
  \item Service 2: The \textit{Registrant A} is configured explicitly.
  \item Service 3: The \textit{Registrant B} is configured explicitly.
\end{itemize}

So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
In the first step each Registrant performs information collection from all assigned services sequentially. Registrants themselves are executing in parallel. For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}). After collection  the aggregated information is sent to the corresponding ISISes independently by each other Registrant (\textit{A5}, \textit{B5}).

\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Registration.png}}}
\caption{\label{fig:registration_overview}Overview of the registration process} }
\end{figure}

This registration operation is done once during the start-up phase and periodically according to (per service) configured periods.

@@

\section{Configuration}
\label{sec:service_configuration}

% The registration operation is done by the Registrants and described in Section \ref{service_registrantion}.

% section configuration (end)
